---
title: "decathlon_data_cleaning_file"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Decathlon Data Cleaning
```{r}
library(tidyverse)
library(here)
library(janitor)
```


# Task 1 - Decathlon Data
```{r}
here::here()
```

## Open file and read it

```{r}
decathlon <- read_rds(here("task_1/raw_data/decathlon.rds"))
```


```{r}
view(decathlon)
```


## 1.1.1 Clean the data  

### One column doesn't have a name. Use rownames_to_column function to give it a name.
```{r}
decathlon_new <- rownames_to_column(decathlon, var = "athlete")
```


### Now clean the variable names to add underscores, take-away capitals etc. using the clean_names function

```{r}
decathlon_save <- decathlon_new
```

 
```{r}
decathlon_clean <- clean_names(decathlon_save, case = "snake") %>% 
  mutate(athlete = str_to_lower(athlete))
  
```


### Let's now lengthen the data using the pivot_longer function and create two new variable names: events and performance. 
### This last part is critical to make the data easier to analyse.
<details>
<summary>**Answer**</summary>
```{r}
decathlon_long <- pivot_longer(decathlon_clean, cols = x100m:x1500m, names_to = "events", values_to = "performance")
```
</details>


## Create a csv with the daz-whiter-than-white data ready to be analysed
```{r}
write_csv(decathlon_long, here("task_1/clean_data/decathlon_long.csv"))
```







