---
title: "decathlon_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(tidyverse)
```


```{r}
library(here)
```

```{r}
here::here()
```


```{r}
decathlon_analysis <- read_csv(here("task_1/clean_data/decathlon_long.csv"))
```

## Who had the longest long jump seen in the data?

```{r}
decathlon_analysis %>% 
  filter(events == "long_jump") %>% 
  arrange(desc(performance))
```
| Athlete    | Competition   | Event     | Performance |
|------------|---------------|-----------|-------------|
| Brian Clay | Olympic Games | Long Jump | 7.96m       |
<br>

## What was the average 100m time in each competition?

```{r}
decathlon_analysis %>% 
  filter(events == "x100m") %>% 
  group_by(competition) %>% 
  summarise(average_100m = round(mean(performance),2))
```
| Competition   | Average 100m time |
|---------------|-------------------|
| Decastar      | 11.18s            |
| Olympic Games | 10.92s            |
<br>

## Who had the highest total points across both competitions?
```{r}
decathlon_analysis %>%
  distinct(athlete, competition, points) %>% 
  group_by(athlete) %>% 
  summarise(total_points = sum(points)) %>% 
  arrange(desc(total_points))
```
| Athlete      | Total Points (across both competitions) |
|--------------|-----------------------------------------|
| Roman Sebrle | 17,110                                  |
<br>

## What was the shot-put scores for the top three competitors in each competition?
```{r}
decathlon_analysis %>%
  distinct(athlete, competition, events, performance) %>% 
  filter(events == "shot_put") %>% 
  filter(competition == "OlympicG") %>% 
  arrange(desc(performance))
```
```{r}
decathlon_analysis %>%
  distinct(athlete, competition, events, performance) %>% 
  filter(events == "shot_put") %>% 
  filter(competition == "Decastar") %>% 
  arrange(desc(performance))
```
| Athlete          | Competion     | Event    | Performance |
|------------------|---------------|----------|-------------|
| Roman Sebrle     | Olympic Games | Shot Put | 16.36m      |
| Dmitriy Karpov   | Olympic Games | Shot Put | 15.93m      |
| Dean Macey       | Olympic Games | Shot Put | 15.73m      |
| Oleksandr Yurkov | Decastar      | Shot Put | 15.19m      |
| Roman Sebrle     | Decastar      | Shot Put | 14.83m      |
| Dmitriy Karpov   | Decastar      | Shot Put | 14.77m      |
<br>

## What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?
```{r}
decathlon_analysis %>%
  distinct(events, performance, points) %>% 
  filter(events == "x400m") %>% 
  arrange(performance)
  
```




























